Dynamical modeling of tidal streams

### 2014-01-01 tex/ms.tex
streams: discussion of stream properties

>>>>

The generative model for a tidal stream proposed here also connects
the dynamics of the stream, in particular its average track and the
dispersion around this track, to the properties of the
progenitor. Primarily, we can constrain $\sigv$, which is proportional
to the velocity dispersion or mass$^{1/3}$ of the progenitor (see
\citealt{Sanders13a}), from the observed width of the stream, even if
the stream's sky projection is the only projection for which we can
resolve the stream. Further $N$-body simulations are necessary to
characterize the exact relation between the \sigv\ parameter and the
velocity dispersion of the progenitor.

Our calculation of the mean stream orbit as a function of angle along
the stream also points the way to determining the location of an
``orphan'' stream's progenitor (an orphan stream is a stream for which
the progenitor is not know). This requires high-quality data that
allows the average orbit as a function of position along the stream to
be determined. If this can be done, then the observation of a steady
change in the mean orbit followed by a plateau indicates that the
progenitor lies in the direction of the plateau. The change in the
mean orbit along the stream is a function of the mean orbit offset
($\Delta \Omega^m$ in the direction of $\Delta \opar$), the dispersion
in orbit offsets ($\sigma^2_{\Omega,1}$ in $\Delta \opar$), and the
disruption time $t_d$. The first two of these can be measured from the
width of the stream and the plateau in the mean orbit as a function of
stream position. The disruption time $t_d$ can then be measured from
the change in mean orbit at the end of the stream. With all of the
parameters of the forward model measured, we can extrapolate the
stream backward to the progenitor position. Similarly, we can
extrapolate a leading stream to its trailing counterpart, if this is
unknown (and vice versa).

While this requires high-quality data, this should be achievable in
the near future. We can average or ``stack'' noisily measured
phase--space coordinates in small segments of the stream to determine
high quality average phase--space positions and use these instead of
data on individual stream stars. These considerations show that
further observations of members of the GD-1 and Orphan streams would
be highly informative for determining the location or fate of their
progenitors.
